---
title: "a3_task4_morganne_sigismonti.Rmd"
author: "Morganne Sigismonti"
date: "11/2/2020"
output: html_document
---
# Assignment 3 Part 4
## Naples Reef (January 2010) Purple Urchins Wrangling and Analysis
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(janitor)
library(here)
library(kableExtra)
```

### Read in the urchins data
```{r, message=FALSE}
urchins_data <- read_csv(here("data", "urchins.csv")) %>% 
  clean_names()
```

## 1) Comparing the mean size of purple urchins in CONTROL (no giant kelp removal) or ANNUAL (annual kelp removal) sections of Naples Reef in Jauary 2010

### a) Wrangling the data
- Here I wrangled the urchins data to only include Naples Reef observations for purple urchins observed in January 2010

```{r, include=FALSE}
  urchins_clean_dates <- urchins_data %>%
  mutate(date = mdy(date))

class(urchins_clean_dates$date)

purple_urchins_comp <- urchins_clean_dates %>%
  mutate(date = ymd(date)) %>%
  mutate(month = month(date)) %>% 
  mutate(year = year(date)) %>% 
  filter(year == 2010,
         month == 1) %>%
  mutate(site_full = recode(site, "NAPL" = "Naples Reef")) %>%
  filter(common_name == "Purple Urchin") %>%
  filter(treatment %in% c("ANNUAL", "CONTROL")) %>%
  filter(site_full == "Naples Reef") %>% 
  select(site_full, common_name, month, year, treatment, size)
  
  
  
 
```

### b) Exploratory graphing to see size distribution for each of the treatments to assess normality

- Here I made a histogram and qqplot

```{r}
ggplot(data = purple_urchins_comp, aes(x = size)) +
  geom_histogram(binwidth = .5) +
  facet_wrap(~ treatment)

ggplot(data = purple_urchins_comp, aes(sample = size)) +
  geom_qq() +
  facet_wrap(~treatment)
```

HELP 1 - 2 sentences describing your main takeaways about the distribution of urchin sizes for each group. HELP


### c) Created a basic table of descriptive statistics for CONTROL and ANNUAL groups

- Here I made a table of both groups with their respective mean, standard deviation, and sample size)

```{r}
purple_urchin_stats <- purple_urchins_comp %>% 
  group_by(treatment) %>% 
  summarize(mean = mean(size),
            sd = sd(size),
            sample_size = n())

kbl(purple_urchin_stats)%>%
   kable_classic()
```

### d) Performed a two-sample t-test to compare means between the two groups

- Here I created vectors containing observations for each sample

```{r}

compact_sample <- compact_suv %>% 
  filter(class == "compact") %>% 
  pull(cty)

suv_sample <- compact_suv %>% 
  filter(class == "suv") %>% 
  pull(cty)
```

### 5. Then run a two-sample t-test with those vectors as the inputs

```{r}
t.test(compact_sample, suv_sample)
```

